---
title: "stats778-casestudy"
output: html_document
date: "2025-04-09"
---
# Load required library
```{r}
library(MASS)  # for mvrnorm
library(quantreg)
library(readr)
library(dplyr)
library(ggplot2)
library(tidyr)
```

# Data generation
```{r}
simulate_cluster <- function(cluster_id, cluster_size, beta, rho = 0.5) {
  p <- length(beta)
  X <- matrix(rnorm(cluster_size * p), ncol = p)
  colnames(X) <- paste0("X", 1:p)
  
  Sigma <- matrix(rho, cluster_size, cluster_size)
  diag(Sigma) <- 1
  error_raw <- mvrnorm(n = 1, mu = rep(0, cluster_size), Sigma = Sigma)
  error_star <- error_raw - median(error_raw)
  
  Y <- X %*% beta + error_star
  data.frame(cluster_id = cluster_id, Y = as.vector(Y), X, error_star)
}

set.seed(123)
n_clusters <- 30
beta_true <- c(0.5, 1, 1.5, 2, -1)
rho <- 0.5

cluster_sizes <- c(20, 20, 20, 20, 100, 20, 100, 100, 20, 20, 
                   100, 100, 100, 100, 20, 100, 20, 20, 100, 100,
                   20, 100, 20, 100, 20, 100, 20, 100, 20, 100)

#  Generate all clusters 
data <- do.call(rbind, lapply(1:n_clusters, function(i) {
  simulate_cluster(i, cluster_sizes[i], beta = beta_true, rho = rho)
}))

write_csv(data, "data.csv")
```

# Quantile regression
```{r}
# Fit baseline quantile regression ---
tau <- 0.5
formula_qr <- Y ~ X1 + X2 + X3 + X4 + X5

# Estimate quantile regression 
model_qr <- rq(formula_qr, data = data, tau = tau)
beta_hat <- coef(model_qr)
```


# Wild gradient bootstrap
```{r}
rMammen <- function(n) {
  p <- (sqrt(5) + 1) / (2 * sqrt(5))
  u <- runif(n)
  ifelse(u < p, -(sqrt(5) - 1)/2, (sqrt(5) + 1)/2)
}

cluster_ids <- unique(data$cluster_id)
n_clusters <- length(cluster_ids)
W <- rMammen(n_clusters)

# Compute bootstrap gradient Wn(tau) 
compute_Wn <- function(data, model, tau, W) {
  grad <- rep(0, length(coef(model)))
  for (i in seq_along(cluster_ids)) {
    cluster_data <- data[data$cluster_id == cluster_ids[i], ]
    psi_vals <- tau - as.numeric(cluster_data$Y < predict(model, newdata = cluster_data))
    X_mat <- model.matrix(model, data = cluster_data)
    grad <- grad + W[i] * colSums(psi_vals * X_mat)
  }
  grad / sqrt(n_clusters)
}

Wn_tau <- compute_Wn(data, model_qr, tau, W)

X_star <- -sqrt(n_clusters) * Wn_tau / tau
Y_star <- max(data$Y) * 10
X_names <- names(beta_hat)[-1]  # exclude intercept
synthetic_row <- data.frame(matrix(X_star[-1], nrow = 1))
colnames(synthetic_row) <- X_names
synthetic_row$Y <- Y_star
synthetic_row$cluster <- max(data$cluster_id) + 1

augmented_data <- bind_rows(data, synthetic_row)
model_star <- rq(formula_qr, data = augmented_data, tau = tau)

wild_gradient_bootstrap <- function(data, model, tau, B = 500) {
  cluster_ids <- unique(data$cluster_id)
  n_clusters <- length(cluster_ids)
  beta_hat <- coef(model)
  results <- matrix(NA, nrow = B, ncol = length(beta_hat))
  colnames(results) <- names(beta_hat)
  
  for (b in 1:B) {
    W <- rMammen(n_clusters)
    Wn_tau <- compute_Wn(data, model, tau, W)
    X_star <- -sqrt(n_clusters) * Wn_tau / tau
    Y_star <- max(data$Y) * 10
    
    X_names <- names(beta_hat)[-1]
    synthetic_row <- data.frame(matrix(X_star[-1], nrow = 1))
    colnames(synthetic_row) <- X_names
    synthetic_row$Y <- Y_star
    synthetic_row$cluster <- max(data$cluster_id) + 1
    
    augmented_data <- bind_rows(data, synthetic_row)
    model_star <- rq(formula_qr, data = augmented_data, tau = tau)
    results[b, ] <- coef(model_star)
  }
  return(results)
}
```



# standard boostrap 
```{r}
# Define a function to compute bootstrap estimates
standard_bootstrap_qr <- function(data, formula_qr, tau = tau, B = B, seed = 123) {
  n <- nrow(data)
  boot_coefs <- matrix(NA, nrow = B, ncol = length(coef(rq(formula_qr, data = data, tau = tau))))
  
  for (b in 1:B) {
    boot_indices <- sample(1:n, size = n, replace = TRUE)
    boot_data <- data[boot_indices, ]
    boot_model <- rq(formula_qr, data = boot_data, tau = tau)
    boot_coefs[b, ] <- coef(boot_model)
  }
  
  colnames(boot_coefs) <- names(coef(boot_model))
  return(boot_coefs)
}
```


# Calculate coverage rate for two methods
```{r}
# caluclate coverage rate for wgb and standard boostrap for single tau and rho value
evaluate_coverage <- function(n_sim = 100, B = 500, tau = 0.5, rho = 0.5,
                              beta_true = c(0.5, 1, 1.5, 2, -1)) {
  formula_qr <- Y ~ X1 + X2 + X3 + X4 + X5
  coef_names <- paste0("X", 1:length(beta_true))
  
  coverage_counts <- matrix(0, nrow = length(beta_true), ncol = 2)
  colnames(coverage_counts) <- c("WGB", "Standard")
  rownames(coverage_counts) <- coef_names
  
  for (sim in 1:n_sim) {
    # Simulate data 
    cluster_sizes <- c(20, 20, 20, 20, 100, 20, 100, 100, 20, 20, 
                       100, 100, 100, 100, 20, 100, 20, 20, 100, 100,
                       20, 100, 20, 100, 20, 100, 20, 100, 20, 100)
    data_co <- do.call(rbind, lapply(1:30, function(i) {
      simulate_cluster(i, cluster_sizes[i], beta_true, rho)
    }))
    
    # Base QR fit 
    formula <- Y ~ X1 + X2 + X3 + X4 + X5
    model <- rq(formula, data = data_co, tau = tau)
    
    # WGB 
    coef_wgb <- wild_gradient_bootstrap(data_co, model, tau, B)
    ci_lower_wgb <- apply(coef_wgb, 2, quantile, 0.025)
    ci_upper_wgb <- apply(coef_wgb, 2, quantile, 0.975)
    
    # Standard Bootstrap 
    coefs_standard <- standard_bootstrap_qr(data_co, model, tau, B)
    ci_lower_standard <- apply(coefs_standard, 2, quantile, probs = 0.025)
    ci_upper_standard <- apply(coefs_standard, 2, quantile, probs = 0.975)
    
    # Check coverage (skip intercept) 
    for (j in 1:length(beta_true)) {
      idx <- j + 1  # skip intercept
      # WGB
      if (beta_true[j] >= ci_lower_wgb[idx] && beta_true[j] <= ci_upper_wgb[idx]) {
        coverage_counts[j, "WGB"] <- coverage_counts[j, "WGB"] + 1
      }
      # Standard
      if (beta_true[j] >= ci_lower_standard[idx] && beta_true[j] <= ci_upper_standard[idx]) {
        coverage_counts[j, "Standard"] <- coverage_counts[j, "Standard"] + 1
      }
    }
  }
  
  # Return empirical coverage rates
  round(coverage_counts / n_sim, 5)
}

# combine coverage rate for wgb and standard boostrap for different tau and rho values
# Define quantile levels and correlation values
taus <- c(0.1, 0.5, 0.9)
rhos <- c(0.1, 0.5, 0.9)

# Storage list for results
coverage_results <- list()

# Loop through all tau and rho combinations
for (tau_val in taus) {
  for (rho_val in rhos) {
    message(paste("Running coverage for tau =", tau_val, ", rho =", rho_val))
    
    # Run evaluation
    result <- evaluate_coverage(n_sim = 100, B = 500, tau = tau_val, rho = rho_val)
    
    # Store with descriptive name
    key <- paste0("tau_", tau_val, "_rho_", rho_val)
    coverage_results[[key]] <- result
  }
}


# Flatten the list into a data frame
coverage_df <- do.call(rbind, lapply(names(coverage_results), function(name) {
  df <- as.data.frame(coverage_results[[name]])
  df$Coefficient <- rownames(df)
  df$Setting <- name
  df
})) %>%
  relocate(Setting, Coefficient)
print(coverage_df)
```

# Plot the coverage rate for two methods.
```{r}
# Split into 3 tables by tau
coverage_tau_01 <- coverage_df %>% filter(grepl("tau_0.1", Setting))
coverage_tau_05 <- coverage_df %>% filter(grepl("tau_0.5", Setting))
coverage_tau_09 <- coverage_df %>% filter(grepl("tau_0.9", Setting))

# plot the coverage rate for two methods across different tau and rho values
coverage_long <- coverage_df %>%
  pivot_longer(cols = c("WGB", "Standard"),
               names_to = "Method", values_to = "Coverage")
plot_coverage_by_tau <- function(tau_value) {
  df_plot <- coverage_long %>%
    filter(grepl(paste0("tau_", tau_value), Setting)) %>%
    filter(!is.na(Coverage))  
  
  ggplot(df_plot, aes(x = Coefficient, y = Coverage, fill = Method)) +
    geom_bar(stat = "identity", position = "dodge") +
    geom_hline(aes(yintercept = 0.95, linetype = "95% Target"), color = "darkred", size = 1) +
    scale_linetype_manual(name = NULL, values = c("95% Target" = "dashed")) +
    coord_cartesian(ylim = c(0.85, 1)) +
    facet_wrap(~ Setting, ncol = 3) +
    labs(
      title = paste("Coverage Rate Comparison at Tau =", tau_value),
      y = "Coverage Rate",
      x = "Coefficient",
      fill = "Method"
    ) +
    theme_minimal() +
    theme(
      plot.title = element_text(size = 14, face = "bold"),
      legend.position = "right"
    )
}


plot_coverage_by_tau("0.1")
plot_coverage_by_tau("0.5")
plot_coverage_by_tau("0.9")
```

# combine calculate the whole tau and rho values for wild gradient bootstrap and standard bootstrap
```{r}
# Quantile levels and correlations
taus <- c(0.1, 0.5, 0.9)
rhos <- c(0.1, 0.5, 0.9)
B     <- 500

# Store results
results_list <- list()

for (tau in taus) {
  for (rho in rhos) {
    # Simulate data with this rho
    set.seed(123)
    data <- do.call(rbind, lapply(1:length(cluster_sizes), function(i) {
      simulate_cluster(i, cluster_sizes[i], beta = beta_true, rho = rho)
    }))
    
    # Fit baseline model
    model_qr <- rq(formula_qr, data = data, tau = tau)
    beta_hat <- coef(model_qr)
    
    # Wild Gradient Bootstrap 
    wgb_coefs <- wild_gradient_bootstrap(data, model_qr, tau, B)
    wgb_bias <- colMeans(wgb_coefs) - beta_true
    wgb_se <- apply(wgb_coefs, 2, sd)
    wgb_ci <- t(apply(wgb_coefs, 2, quantile, probs = c(0.025, 0.975)))
    
    wgb_df <- data.frame(
      Method = "WGB",
      tau = tau,
      rho = rho,
      Coefficient = names(beta_hat),
      Estimate = beta_hat,
      Bias = wgb_bias,
      SE = wgb_se,
      CI_width = wgb_ci[,2] - wgb_ci[,1],
      CI_lower = wgb_ci[,1],
      CI_upper = wgb_ci[,2]
    )
    
    # Standard Bootstrap 
    standard_coefs <- standard_bootstrap_qr(data, model_qr, tau, B)
    standard_bias <- colMeans(standard_coefs) - beta_true
    standard_se <- apply(standard_coefs, 2, sd)
    standard_ci <- t(apply(standard_coefs, 2, quantile, probs = c(0.025, 0.975)))
    
    standard_df <- data.frame(
      Method = "Standard",
      tau = tau,
      rho = rho,
      Coefficient = names(beta_hat),
      Estimate = beta_hat,
      Bias = standard_bias,
      SE = standard_se,
      CI_width = standard_ci[,2] - standard_ci[,1],
      CI_lower = standard_ci[,1],
      CI_upper = standard_ci[,2]
    )
    
    # Combine and store
    results_list[[paste0("tau_", tau, "_rho_", rho)]] <- rbind(wgb_df, standard_df)
  }
}
# Combine all into one data frame
final_results <- do.call(rbind, results_list)
print(final_results)
```


# figure bias for two methods across different tau and rho values
```{r}
for (tau_val in unique(final_results$tau)) {
  df <- subset(final_results, tau == tau_val)
  
  p <- ggplot(df, aes(x = rho, y = Bias, color = Method, group = Method)) +
    geom_line(linewidth = 1) +
    geom_point(size = 2) +
    facet_wrap(~ Coefficient, nrow = 2, scales = "free_y") +
    labs(title = paste("Bias Comparison across rho at tau =", tau_val),
         x = "Intra-cluster correlation (rho)", y = "Bias") +
    scale_x_continuous(breaks = sort(unique(df$rho))) +
    theme_minimal() +
    theme(strip.text = element_text(size = 10, face = "bold"))
  
  print(p)
}
```


# figure SE for two methods across different tau and rho values
```{r}
for (tau_val in unique(final_results$tau)) {
  df <- subset(final_results, tau == tau_val)
  
  p <- ggplot(df, aes(x = rho, y = SE, color = Method, group = Method)) +
    geom_line(linewidth = 1) +
    geom_point(size = 2) +
    facet_wrap(~ Coefficient, nrow = 2, scales = "free_y") +
    labs(title = paste("SE Comparison across rho at tau =", tau_val),
         x = "Intra-cluster correlation (rho)", y = "Standard Error") +
    scale_x_continuous(breaks = sort(unique(df$rho))) +
    theme_minimal() +
    theme(strip.text = element_text(size = 10, face = "bold"))
  
  print(p)
}
```


# figure CI width for two methods across different tau and rho values
```{r}
for (tau_val in unique(final_results$tau)) {
  df <- subset(final_results, tau == tau_val)
  
  p <- ggplot(df, aes(x = rho, y = CI_width, color = Method, group = Method)) +
    geom_line(linewidth = 1) +
    geom_point(size = 2) +
    facet_wrap(~ Coefficient, nrow = 2, scales = "free_y") +
    labs(title = paste("CI_width Comparison across rho at tau =", tau_val),
         x = "Intra-cluster correlation (rho)", y = "CI_width") +
    scale_x_continuous(breaks = sort(unique(df$rho))) +
    theme_minimal() +
    theme(strip.text = element_text(size = 10, face = "bold"))
  
  print(p)
}
```

